---
title: "Project 3: myfirstpackage Tutorial"
author: "Matteo Vindrola"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{myfirstpackage Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(myfirstpackage)
```
A brief introductory paragraph explaining the package
This paragraph should include instructions for how to install your package from Github.
The install instructions should be demonstrated with code that is not evaluated (eval = FALSE).
Make sure to include a call to library() for your package.

A tutorial for my_t.test
Use the lifeExp data from my_gapminder.
Demonstrate a test of the hypothesis
H0:μHa:μ=60,≠60.
Demonstrate a test of the hypothesis
H0:μHa:μ=60,<60.
Demonstrate a test of the hypothesis
H0:μHa:μ=60,>60.
For each of the tests above, carefully interpret the results using a p-value cut-off of α=0.05.

A tutorial for my_lm
Demonstrate a regression using lifeExp as your response variable and gdpPercap and continent as explanatory variables.
Carefully interpret the gdpPercap coefficient.
Write the hypothesis test associated with the gdpPercap coefficient.
Carefully interpret the results the gdpPercap hypothesis test using a p-value cut-off of α=0.05.
Use ggplot2 to plot the Actual vs. Fitted values.
Interpret the Actual vs. Fitted plot and make a statement on what it tells you about model fit.

A tutorial for my_knn_cv
Predict output class continent using covariates gdpPercap and lifeExp.
Use 5-fold cross validation (k_cv = 5).
Iterate from k_nn=1,…,10:
For each value of k_nn, record the training misclassification rate and the CV misclassification rate (output from your function).
State which model you would choose based on the training misclassification rates and which model you would choose based on the CV misclassification rates.
Discuss which model you would choose in practice and why.
Your explanation(s) should include a general description of what the process of cross-validation is doing and why it is useful.

A tutorial for my_rf_cv
From Lab 4, switch the input data from iris to my_gapminder.
Predict lifeExp using covariate gdpPercap.
Iterate through k in c(2, 5, 10):
For each value of k, run your function 30 times.
For each of the 30 iterations, store the CV estimated MSE.
Use ggplot2 with 3 boxplots to display these data in an informative way. There should be a boxplot associated with each value of k, representing 30 simulations each.
Use a table to display the average CV estimate and the standard deviation of the CV estimates across k. Your table should have 3 rows (one for each value of k) and 2 columns (one for the mean and the other for the standard deviation).
Discuss the results you observe in the boxplots and table. Compare the means and standard deviations of the the different values of k and comment on why you think this is the case.
